---
title: "Basic Data cleaning and export"
author: "Anuradha Vyas"
date: "2024-01-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r, warning=FALSE, message= FALSE}
library(writexl) #to export excel sheet
library(dplyr)
library(openxlsx)
library(readxl) #to read excel
```

# Create and Print Dataset
```{r}
Original_Data <- data.frame(
  A = c(1, 2, 3, 4, 5),
  B = c("Red", "Blue", "Green", "Yellow", "Red"),
  C = c(10.5, 15.2, 8.7, 12.0, 9.3),
  D = c(TRUE, FALSE, TRUE, FALSE, TRUE),
  E = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
)

# Print the new dataset
print(Original_Data)
```

# Export Data to Excel
```{r}
File_path <- "C:/Users/smile/Desktop/Original_data_1.xlsx"
write_xlsx(Original_Data, File_path)
```

# Rename Columns
```{r}
Original_Data_rename <- rename(Original_Data, id = A, color = B, 
                        age = C, T_F = D, 
                        Days = E)
Original_Data_rename
```

# Count Data by Category

```{r}
count(Original_Data_rename, id) #id is not converted to categorical
count(Original_Data_rename, T_F)  # Count the occurrences of 'T_F'
count(Original_Data_rename, T_F, color) # Count the occurrences of unique combinations of 'T_F' and 'color'
```

# Simplify Color Column
```{r}

Original_Data_rename <- mutate(Original_Data_rename, color.simplified = case_when(
  color == "Red" ~ "1",
  color == "Blue" ~ "2",
  color == "Green" ~ "3",
  color == "Yellow" ~ "4"
))

select(Original_Data_rename, c(color, color.simplified))
```

# Calculate Mean and SD by Group
```{r}
Original_Data_rename

Original_Data_rename |> 
    mutate(midschool.school.complete = case_when(
      age< 11 ~ "No_midschool",
      age >11 ~ "Midschool")) |> 
    group_by(midschool.school.complete) |> 
    summarize(age.mean = mean(age), age.sd= sd(age))

#count(Original_Data_rename, midschool.school.complete)
```


# Add New Data to Excel file
```{r}
#create new data
New_data <- Original_Data |>select(A, B, C)
print(New_data)
# Load the existing workbook
WB <- loadWorkbook(File_path)

addWorksheet(WB, sheetName = "New_data")
writeData(WB, sheet = "New_data", x = New_data)
```


# Add Additional Data to new sheet in same excel file
```{r}
#create new data 2
Additional_data <- Original_Data |>select(A, D, E)
print(Additional_data)


# Add the new_data to a new sheet in the existing workbook
addWorksheet(WB, sheetName = "Additional_data")
writeData(WB, sheet = "Additional_data", x = Additional_data)

# Save the modified workbook
saveWorkbook(WB, File_path, overwrite = TRUE)
```

# Remove Columns and Combine Data
```{r}
New_data_remove <- Original_Data |>select(-A, -B,-C)
print(New_data_remove)
```

#Edit Data and Save to new sheet in same excel file
```{r}
# Add the new_data to a new sheet in the existing workbook

addWorksheet(WB, sheetName = "combined_data_NA")
writeData(WB, sheet = "combined_data_NA", x = cbind(Additional_data, New_data))

# Save the modified workbook
saveWorkbook(WB, File_path, overwrite = TRUE)
```



```{r}
Edited_data <- Original_Data[-c(3,4),]
print(Edited_data)


addWorksheet(WB, sheetName = "Edited_data")
writeData(WB, sheet = "Edited_data", x = Edited_data)

# Save the modified workbook
saveWorkbook(WB, File_path, overwrite = TRUE)
```


#Read Data from different Sheets in same excel file

```{r}
#read data sheet 1
New_data_1 <- read_xlsx("C:/Users/smile/Desktop/Original_data_1.xlsx", sheet ="New_data")
print(New_data_1)
```

```{r}
#read data sheet 2
Additional_data_1 <- read_xlsx("C:/Users/smile/Desktop/Original_data_1.xlsx", sheet ="Additional_data")
print(Additional_data_1)
```


```{r}
#read data sheet 1
combined_data_NA_1 <- read_xlsx("C:/Users/smile/Desktop/Original_data_1.xlsx", sheet ="combined_data_NA")
print(combined_data_NA_1)
```

